Malacology: A Programmable Storage System

### 2016-10-18 paper/paper.tex
paper updates
Signed-off-by: Noah Watkins <noahwatkins@gmail.com>

A number of recent developments have disrupted traditional storage systems.
First, the falling prices of flash storage and the availability of new types of
non-volatile memory that are orders of magnitude faster than traditional
spinning media are moving overall performance bottlenecks away from storage
devices to CPUs and networking, and pressure storage systems to shorten their
code paths and incorporate new
optimizations~\cite{gray_tape_2007,gray_flash_2008}. 
Second, the demand for
managing structured data and flexible consistency semantics at scale pressure
big data processing systems to use storage abstractions that can meet these
demands~\cite{apache_contributors_parquet_2014}. 
\paa{the previous sentence doesn't scan well. do we just want to say that 
emerging ``big data'' applications demand interface evolution to support flexible
consistency as well as flexible structured data representation?}
Finally, production-quality
scalable storage systems available as open source software have established and
are continuing to establish new, \emph{de-facto} API standards at a faster pace
than traditional standards
bodies~\cite{snia_implementing_2014,linux_foundation_kinetic_2015}.

@@

The main challenge of designing a programmable storage system is choosing the
right internal abstractions and picking the correct layers for exposing them.
In this paper, we do not present an exhaustive list of possible internal
abstractions nor do we contend that the abstractions we choose provide the best
trade-offs for all applications. For example, if consensus is correctly
exposed, one could implement versioning, serialization, or capabilities on top;
but perhaps versioning, serialization, and capabilities would be better
interfaces for a given set of applications. These questions are not answered in
this paper and instead we focus on showing the feasibility of building such a
system, given advances in the quality and robustness of today's storage stacks.

@@

The default behavior of the monitor can be seen as a Paxos-based notification
system, similar to the one introduced in~\cite{burrows_chubby_2006}, allowing
clients to identify when new values (termed epochs in Ceph) are associated to
given maps. While Ceph does not expose this service directly, it does allow
limited use by clients through a key-value service intended for low-volume,
small size read-only configuration information. While a key-value service is
generally useful, it doesn't provide many of the useful services hidden within
the monitoring framework that would be required for applications with more
demanding requirements, such as creating arbitrary maps or associating
application-specific logic that can be executed for particular maps (or values
in a map).  Since the monitor is intended to be out of high-performance I/O
paths, a general guideline is to make use of this functionality infrequently
and to assign small values to maps.

>>>>

A number of recent developments have disrupted traditional storage systems.
First, the falling prices of flash storage and the availability of new types of
non-volatile memory that are orders of magnitude faster than traditional
spinning media are moving overall performance bottlenecks away from storage
devices to CPUs and networking, and pressure storage systems to shorten their
code paths and incorporate new
optimizations~\cite{gray_tape_2007,gray_flash_2008}. 
Second, emerging ``big data'' applications demand interface evolution to
support flexible consistency as well as flexible structured data
representations.  ~\cite{apache_contributors_parquet_2014}. 
Finally, production-quality
scalable storage systems available as open source software have established and
are continuing to establish new, \emph{de-facto} API standards at a faster pace
than traditional standards
bodies~\cite{snia_implementing_2014,linux_foundation_kinetic_2015}.

@@

The main challenge of designing a programmable storage system is choosing the
right internal abstractions and picking the correct layers for exposing them.
In this paper, we do not present an exhaustive list of possible internal
abstractions nor do we contend that the abstractions we choose provide the best
trade-offs for all applications.
For example, if consensus is correctly exposed one could implement
high-level features like versioning, serialization, or various flavors
of strongly consistent data management on top; but perhaps a low-level
consensus interface is suited well for a particular set of applications.
These questions are not answered in
this paper and instead we focus on showing the feasibility of building such a
system, given advances in the quality and robustness of today's storage stacks.

@@

The default behavior of the monitor can be seen as a Paxos-based notification
system, similar to the one introduced in~\cite{burrows_chubby_2006}, allowing
clients to identify when new values (termed epochs in Ceph) are associated to
given maps.  While Ceph does not expose this service directly, a key-value
service designed for managing configuration metadata that is built on top of
the consensus engine is an example of a high-level service within Ceph that
re-uses existing abstractions.  While a key-value service is generally useful,
it doesn't provide many of the useful services hidden within the monitoring
framework that would be required for applications with more demanding
requirements, such as creating arbitrary maps or associating
application-specific logic that can be executed for particular maps (or values
in a map).  Since the monitor is intended to be out of high-performance I/O
paths, a general guideline is to make use of this functionality infrequently
and to assign small values to maps.
